---
title: "Updated code"
author: "Keith Welch and Taylor Meyers"
date: "12/15/2016"
output: html_document
---

Read in data and create a dataset that removes observations missing education data
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr) 
library(forcats)
library(foreign)
library(haven)
library(lubridate)
library(broom)
library(ISLR)
library(modelr)  
library(ROCR)
library(randomForest)


workforce_full <- read_dta("full_workforce_file.dta")

# sort data frame by employee id and fiscal year and filter out unnecessary data
workforce <- workforce_full %>%
  arrange(jobemplid, fiscal_year) %>%
  filter(fiscal_year >2001 & fiscal_year < 2017) %>%
  mutate(fiscal_year = as.integer(fiscal_year)) %>%
  group_by(jobemplid) %>%
  mutate(missing_data = as.numeric(language_descr=="")) %>%
  filter(missing_data != 1)

workforce <- workforce %>%
  ungroup() %>%
  mutate(language_descr = ifelse(language_descr == "", NULL, language_descr),
         language_descr = factor(language_descr))


```

Create degree achieved dummy variables
```{r education experience}

workforce <- workforce %>%
  mutate(bachelor_temp = as.numeric(str_detect(degreedescr, "Bachelor")),
         master_temp = as.numeric(str_detect(degreedescr, "Master"))) %>%
  group_by(jobemplid) %>%
  mutate(bachelor = max(bachelor_temp),
         master = max(master_temp))

```

Create language variables
```{r language variables}
# create number of languages spoken variable and dummies
workforce <- workforce %>%
  group_by(jobemplid) %>%
  mutate(number_languages = n_distinct(language_descr))

workforce <- workforce %>%
  mutate(one_language = as.numeric(number_languages == 1),
         two_languages = as.numeric(number_languages == 2),
         three_language = as.numeric(number_languages == 3),
         four_above_language = as.numeric(number_languages > 3))

# function to create language dummies
create_language_dummy <- function(language, data) {

  temp <- data %>%
  group_by(jobemplid) %>%
  mutate(language = max(as.numeric(language_descr == language)))

  return(temp$language)

}
languages <- levels(workforce$language_descr)

language_dummies <- sapply(languages, create_language_dummy, 
                           data = workforce)

language_dummies <- tbl_df(language_dummies)

workforce <- bind_cols(workforce, language_dummies)

# rename NULL and Sign Language variables (assuming that NULL indicates English Only)
workforce$English_Only <- workforce$'NULL'
workforce$'NULL' <- NULL
workforce$Sign_Language <- workforce$'Sign Language'
workforce$'Sign Language' <- NULL


```

Create a variable to identify teachers and variables related to job description
```{r teacher variable}
workforce <- workforce %>%
  mutate(teacher_obs = as.numeric(jobdescr == "Adapted PE Teacher"
                              | jobdescr == "Bilingual Classroom Teacher"
                              | jobdescr == "Dept Head / Classroom Teacher"
                              | jobdescr == "ESL Teacher"
                              | jobdescr == "ESL/Bilingual Teacher"
                              | jobdescr == "Eld Classroom Teacher"
                              | jobdescr == "Homebound / Hospital Svcs Tchr"
                              | jobdescr == "Itinerant Dance Teacher"
                              | jobdescr == "Itinerant Drama Teacher"
                              | jobdescr == "Itinerant Music Teacher"
                              | jobdescr == "Itinerant Visual Art Teacher"
                              | jobdescr == "Mild/Mod RS/SDC (No Autism)"
                              | jobdescr == "Mild/Mod Tchr SDC Only"
                              | jobdescr == "Mild/Mod Tchr SDC Only (No Au)"
                              | jobdescr == "Mild/Moderate RS/SDC"
                              | jobdescr == "Moderate/Severe RS/SDC"
                              | jobdescr == "Orientation Mobility Tch Only"
                              | jobdescr == "Orthopedically Impaired RS/SDC"
                              | jobdescr == "Peer Resource Teacher"
                              | jobdescr == "Prop \"A \"Sub Teacher"
                              | jobdescr == "ROTC Teacher"
                              | jobdescr == "Regular Classroom Teacher"
                              | jobdescr == "Resource Teacher"
                              | jobdescr == "Secondary Art Teacher"
                              | jobdescr == "Secondary Music Teacher"
                              | jobdescr == "TSA - Physical Education"
                              | jobdescr == "Teacher Sab Leave"
                              | jobdescr == "Teacher Special Assign"
                              | jobdescr == "Teacher/Resource Teacher"
                              | jobdescr == "Technical Resource Tch"
                              | jobdescr == "Transitional KG Teacher"
                              | jobdescr == "Vis. Impr. & Orient. Mob. Tchr"
                              | jobdescr == "Visual Impairment Tchr"))

# generate a teacher variable that shows if any of a teacher's observations for a given fiscal year are a teacher job description
workforce <- workforce %>%
  group_by(jobemplid, fiscal_year) %>%
  mutate(teacher = max(teacher_obs)) %>%
  arrange(jobemplid, fiscal_year) 


# generate dummy variable that is equal to one if the employee was ever a teacher
workforce <- workforce %>%
  group_by(jobemplid) %>%
  mutate(ever_teacher = max(teacher_obs))

# generate variable that counts how many job descriptions teachers have each year and corresponding dummies
workforce <- workforce %>%
  group_by(jobemplid, fiscal_year) %>%
  mutate(n_job_descriptions = length(unique((jobdescr[!is.na(jobdescr)]))))

workforce <- workforce %>%
  mutate(one_jobdesc = as.numeric(n_job_descriptions == 1),
         two_jobdesc = as.numeric(n_job_descriptions == 2),
         three_jobdesc = as.numeric(n_job_descriptions == 3),
         four_above_jobdesc = as.numeric(n_job_descriptions > 3))


```

Create variables related to employee license type
```{r license type}
# create a license type variable
workforce <- workforce %>%
  mutate(license_temp = if_else(teacher_obs == 1, license_type_descr, NULL))

# Variable that counts how many licenses teachers have in each fiscal year and corresponding dummies
workforce <- workforce %>%
  group_by(jobemplid, fiscal_year) %>%
  mutate(n_licenses = length(unique((license_temp[!is.na(license_temp)]))))

workforce <- workforce %>%
  mutate(one_n_licenses = as.numeric(n_licenses == 1),
         two_n_licenses = as.numeric(n_licenses == 2),
         three_n_licenses = as.numeric(n_licenses == 3),
         four_n_licenses = as.numeric(n_licenses == 4),
         five_above_n_licenses = as.numeric(n_licenses > 4))
```

Create a variable identifying full time employees
```{r FTE dummy}
workforce <- workforce %>%
  mutate(FTE_dummy = as.numeric(fte >= 1))
```

Create a variable identifying teachers with emergency or provisional credentials
```{r emergency credential variable}
# # create a dummy variable that is equal to 1 for all observations in each year that a teacher has an emergency or provisional credential (unless the teacher also has a clear credential in that same year
workforce <- workforce %>%
  group_by(jobemplid, fiscal_year) %>%
  mutate(prov_emerg_cred = max(as.numeric(str_detect(credential_term, "Provisional") |
                                               str_detect(credential_term, "Emergency"))))
```

Transform dataset to panel format
```{r transform dataset to panel format}
workforce_panel <- workforce %>%
  group_by(jobemplid, fiscal_year) %>%
  distinct(.keep_all = TRUE) %>%
  filter(ever_teacher == 1)
```

Create variables that identify whether a teacher left the district
```{r left district variables}
# create variables to indicate whether a teacher left the district or switched to a non-teaching job
workforce_panel <- workforce_panel %>%
  ungroup() %>%
  mutate(fiscal_year = as.integer(fiscal_year)) %>%
  mutate(last_entry = as.numeric(
    lead(jobemplid) != jobemplid)) %>%
  mutate(left_district = as.numeric((last_entry == 1 & fiscal_year != 2016) | 
                                      ((((lead(fiscal_year) - fiscal_year)) > 1) & 
                                         last_entry == !1 & 
                                         (lead(jobemplid) == jobemplid)))) %>%
  mutate(changed_to_non_teaching_job = as.numeric(
    ((lead(teacher) - teacher) == -1) &
      (left_district != 1 &
      lead(jobemplid) == jobemplid)))
      
      
# create a dummy variable indicating whether a teacher went on a "leave of absence" if they left the district and then returned 
workforce_panel <- workforce_panel %>%
  group_by(jobemplid) %>%
  mutate(leave_absence = as.numeric(left_district - last_entry == 1))


```

Create variables that identify whether a teacher resigned from the district 
```{r resignation dummy}
workforce_panel <- workforce_panel %>%
   mutate(resignation_temp = as.numeric(action_reasondescr == "Resignation" |
                  action_reasondescr == "Relocation" |
                  action_reasondescr == "Personal Reasons" |
                  action_reasondescr == "Resign Employed by Other  Distr" |
                  action_reasondescr == "Failure to Return from Leave" |
                  action_reasondescr == "San Francisco City & County" |
                  action_reasondescr == "Dissatisfied w/Work Conditions" |
                  action_reasondescr == "Return to School" | 
                  action_reasondescr == "Resign Change Of Profession" |
                  action_reasondescr == "Resign Quit without Cause" |
                  action_reasondescr == "Job Abandonment" |
                  action_reasondescr == "Resign Care Of Children" |
                  action_reasondescr == "Resign Non Reelect" |
                  action_reasondescr == "Resign Dissatisfy Work Conditi" |
                  action_reasondescr == "Resign Teaching Out Of USA" |
                  action_reasondescr == "Health Reasons" |
                  action_reasondescr == "Rsgn Stlmnt Agmt Certs" |
                  action_reasondescr == "Cost-of Living" |
                  action_reasondescr == "Dissatisfied with Pay" |
                  action_reasondescr == "Family Reasons" |
                  action_reasondescr == "Resign Personal Reasons" |
                  action_reasondescr == "Rsgn Per Settlement Agmt" |
                  action_reasondescr == "Illness in Family" |
                  action_reasondescr == "Resign Joining Spouse" |
                  action_reasondescr == "Child/House Care" |
                  action_reasondescr == "Dissatisfied w/Type of Work" |
                  action_reasondescr == "Conversion Retirement" |
                  action_reasondescr == "Dissatisfied w/Comp. Policies" |
                  action_reasondescr == "Marriage" |
                  action_reasondescr == "Resign Maternity" |
                  action_reasondescr == "Resignation-Other Position" |
                  action_reasondescr == "Transfer to City & County" |
                  action_reasondescr == "Transportation Problems"),
          resignation = as.numeric(resignation_temp == 1 & left_district == 1))

# dummy to indicate whether a teacher ever resigned
workforce_panel <- workforce_panel %>%
  group_by(jobemplid) %>%
  mutate(ever_resigned = max(resignation))

```

Create a non-school location variable
```{r}
## create vector of all schools that appear 30 times or less in panel to create an "other_location" variable
workforce_panel <- workforce_panel %>%
  ungroup() %>%
  mutate(locationdescr = factor(locationdescr))

other_locations <- workforce_panel %>%
  group_by(locationdescr) %>%
  summarise(number = sum(ever_teacher)) %>%
  filter(number <= 30)

other_location_names <- as.character(other_locations$locationdescr)

## assign all non-school locations to "non-school location" as locationdescr

non_school_loc_names <- c("Substitute Teachers",
"Special Education Services" ,
"Human Resources" ,
"Summer School Department" ,
"CAO - Curriculum Improvement" ,
"Special Education Sub Paras" ,
"Childrens Centers Substitutes" ,
"Multilingual Pathways Dept" ,
"Physical Education Department" ,
"C&I Teaching & Learning" ,
"Instructional Reform Network" ,
"Elementary Schools" ,
"CAO - Library Service" ,
"CAREER TECHNICAL EDUCATION" ,
"MUSIC - VISUAL & PERFORMING AR" ,
"SFCSD - Pupil Services" ,
"SFCSD-Cnseling & Social Servic" ,
"Childrens Ctr. Administration" ,
"SFCSD - Prevention" ,
"Assessment & Accountability" ,
"Teacher Support & Development" ,
"Superintendent Of Schools" ,
"Multiple Work Locations" ,
"Superintendentâs Zone-Bayview" ,
"ELEM. INSTRUCTIONAL SUP. & OPS" ,
"State & Federally-Funded Proje" ,
"Middle Schools" ,
"Chinese Education Center" ,
"Spec Ed. Id And Assessment" ,
"Gifted & Talented Edu. Office")
  

workforce_panel <- workforce_panel %>%
  mutate(locationdescr = factor(locationdescr),
         locationdescr_new = fct_collapse(locationdescr,
                                          other_location = other_location_names,
                                          non_school_location = non_school_loc_names))
```

Merge 2016 rental data by zip code
```{r merge rental price data}
# merge data (note that the data is from 2016)
hud_rental_data <- read_csv("HUD_data.csv")
hud_rental_data$postal <- hud_rental_data$POSTAL
hud_rental_data$POSTAL <- NULL

workforce_panel <- workforce_panel %>%
  mutate(postal = as.integer(postal))

workforce_panel <- left_join(workforce_panel, hud_rental_data, by = "postal")

# dummy variable indicating whether 2016 median rent was less than $1500
workforce_panel <- workforce_panel %>%
  mutate(rent_0_1500 = as.numeric(median_rent < 1500))

```

Create home zip codes dummy variables
```{r zip code dummies}
# create dummy variables for the most frequently appearing zip codes (home address)
## extract most common zip codes (all that appear 100 times or more in the workforce_panel set)
zip_codes <- workforce_panel %>%
  group_by(postal) %>%
  summarize(number = sum(ever_teacher)) %>%
  filter(number > 99 & !is.na(postal)) %>%
  mutate(postal = factor(postal))

common_zips <- levels(zip_codes$postal)
common_zip_names <- paste(common_zips, "_home_zip", sep="")

## function to create zip code dummies
create_zip_dummy <- function(zip, data){
  
  temp <- data %>%
    mutate(zip_code = as.numeric(postal == zip))
  
  return(temp$zip_code)
}

zip_dummies <- sapply(common_zips, create_zip_dummy, data = workforce_panel)

zip_dummies <- tbl_df(zip_dummies)
names(zip_dummies) <- common_zip_names

workforce_panel <- bind_cols(workforce_panel, zip_dummies)

workforce_panel <- workforce_panel %>%
  mutate(other_home_zip = as.numeric('94014_home_zip' == 0 & 
                                       '94015_home_zip' == 0 & 
                                       '94044_home_zip' == 0 & 
                                       '94066_home_zip' == 0 & 
                                       '94080_home_zip' == 0 & 
                                       '94102_home_zip' == 0 & 
                                       '94103_home_zip' == 0 & 
                                       '94107_home_zip' == 0 & 
                                       '94109_home_zip' == 0 &
                                       '94110_home_zip' == 0 & 
                                       '94112_home_zip' == 0 & 
                                       '94114_home_zip' == 0 & 
                                       '94115_home_zip' == 0 & 
                                       '94116_home_zip' == 0 & 
                                       '94117_home_zip' == 0 & 
                                       '94118_home_zip' == 0 & 
                                       '94121_home_zip' == 0 & 
                                       '94122_home_zip' == 0 &
                                       '94123_home_zip' == 0 & 
                                       '94124_home_zip' == 0 & 
                                       '94127_home_zip' == 0 & 
                                       '94131_home_zip' == 0 & 
                                       '94132_home_zip' == 0 & 
                                       '94133_home_zip' == 0 & 
                                       '94134_home_zip' == 0 & 
                                       '94608_home_zip' == 0 & 
                                       '94609_home_zip' == 0 &
                                       '94610_home_zip' == 0 & 
                                       '94611_home_zip' == 0 & 
                                       '94941_home_zip' == 0))

```

Create variables counting the number of years the teacher has taught in the district
```{r years in district}

workforce_panel <- workforce_panel %>%
  mutate(temp = 1) %>%
  group_by(jobemplid) %>%
  mutate(district_years = cumsum(temp)) 

workforce_panel <- workforce_panel %>%
  mutate(one_district_year = as.numeric(district_years == 1),
         two_district_year = as.numeric(district_years == 2),
         three_district_year = as.numeric(district_years == 3),
         four_district_year = as.numeric(district_years == 4),
         five_above_district_year = as.numeric(district_years > 4)) %>%
  mutate(above_3_district_year = as.numeric(district_years > 3))
```

Create teacher age variables
```{r teacher age}
workforce_panel$birth_year <- year(workforce_panel$birthdate)

workforce_panel$age <- workforce_panel$fiscal_year - workforce_panel$birth_year

workforce_panel <- workforce_panel %>%
  mutate(age_under_30 = as.numeric(age < 30),
         age_30_40 = as.numeric(age >= 30 & age < 40),
         age_over_40 = as.numeric(age > 30))

```

Create year of entry into district variable
```{r year of entry into district}
workforce_panel <- workforce_panel %>%
  ungroup() %>%
  mutate(fiscal_year = as.integer(fiscal_year)) %>%
  group_by(jobemplid) %>%
  mutate(district_entry_date = min(fiscal_year)) %>%
  mutate(district_entry_date = as.factor(district_entry_date))
  
```

Create a variable indicating whether a teacher moved
```{r whether teacher changed home addresses}
workforce_panel <- workforce_panel %>%
  ungroup() %>%
  mutate(moved_house = as.numeric(
    lag(address1) != address1) & (lag(jobemplid) == jobemplid))
```

Create a variable indicating whether a teacher is male
```{r sex}
workforce_panel <- workforce_panel %>%
  mutate(male = as.numeric(sex == "M"))
```

Create race variables
```{r race and ethnicity}
#Race
workforce_panel$xlatlongname <- as.factor(workforce_panel$xlatlongname)
races <- levels(workforce_panel$xlatlongname)


create_race_dummy <- function(race, data) {


  temp <- data %>%
  group_by(jobemplid) %>%
  mutate(race = as.numeric(xlatlongname == race))


  return(temp$race)


}


race_dummies <- sapply(races, create_race_dummy, 
                           data = workforce_panel)


race_dummies <- tbl_df(race_dummies)


workforce_panel <- bind_cols(workforce_panel, race_dummies)

# do some cleaning
workforce_panel$Japanese <- NULL
workforce_panel$Korean <- NULL
workforce_panel$Samoan <- NULL
workforce_panel$Vietnamese <- NULL

```

Create panel dataset for analysis
```{r create workforce panel set for analysis}
workforce_panel_analysis <- workforce_panel %>%
  ungroup() %>%
  mutate(locationdescr = factor(locationdescr),
         postal = factor(postal),
         factor_year = factor(fiscal_year),
         jobemplid = factor(jobemplid)) %>%
  select(-c(all_emplid,  first_name,  last_name,  applicant_id,  college_id,  collegedescr,  degree_id,  degreedescr,  gpa_overall,  major_id,  major,  teaching_degree,  total_hours,  expiration_date,  license_college_id,  licensecollege,  license_state,  license_status_id,  license_status_descr,  license_type_id,  license_type_descr,  license_type_text,  app_language_id,  language_id,  language_descr,  submitted_date,  submittedschoolyear,  effdate,  name,  job_entry_dt,  dept_entry_dt,  empl_rcd,  grade,  step,  jobcode,  jobdescr,  location,  locationdescr,  birthdate,  sex,  sfu_ethnicity,  empl_status,  action_reason,  action_reasondescr,  union_cd,  unionname,  fte,  address1,  address2,  city,  state,  postal,  uccal_emplid,  ethnic_group,  xlatlongname,  primary_indicator,  dist_pct,  credential_term,  valid_start,  valid_end,  unique_employee,  missing_data,  bachelor_temp,  master_temp,  number_languages,  ever_teacher,  n_job_descriptions,  license_temp,  n_licenses,  left_district,  resignation_temp,  ever_resigned,  temp,  district_years)) %>%
  mutate(last_entry_temp = if_else(last_entry == 1, fiscal_year, NULL)) %>%
  group_by(jobemplid) %>%
  mutate(last_entry_grouped = max(last_entry_temp, na.rm = TRUE))

training_set <- workforce_panel_analysis %>%
  filter(last_entry_grouped > 2001 & last_entry_grouped < 2013)

test_set <- workforce_panel_analysis %>%
  filter(last_entry_grouped >= 2013 & last_entry_grouped < 2015) 

holdout_set <- workforce_panel_analysis %>%
  filter(last_entry_grouped == 2015) 


# school_zipcode, studentcount,  type,  percentELL,  percentFRPM,  percentSPED, bachelor_degree_temp,  education_degree_temp,  education_degree,  master_degree_temp,  master_education_degree_temp,  bachelor_gpa_temp,  bachelor_gpa,  master_gpa_temp,  master_gpa,  bachelor_hours_temp,  bachelor_hours,  master_hours_temp,  master_hours

```

Modeling
```{r}
        
# # Step 1 Define null
# null=glm(resignation~1,data=training_set, family = binomial)
# 
# 
# 
# 
# # Step 2 Define full
# full <- glm(resignation ~ bachelor + master + one_language + two_languages + three_language + four_above_language +
#               teacher + one_jobdesc + two_jobdesc + three_jobdesc + four_above_jobdesc + one_n_licenses + two_n_licenses +
#               three_n_licenses + four_n_licenses + five_above_n_licenses + FTE_dummy + one_district_year +
#               two_district_year + three_district_year + four_district_year + five_above_district_year +
#               male + `White/Caucasian` + `African-American` +
#               `American Indian/Native Alaskan` + `Asian Indian` + `Cambodian` + `Chinese` + `Decline To State` +
#               `Filipino` + `Guamanian` + `Japanese` + `Korean` + `Other Asian` + `Other Pacific Islander` + 
#               `Samoan` + `94015_home_zip` + `94102_home_zip` + `94103_home_zip` + `94107_home_zip` + `94109_home_zip` +
#               `94110_home_zip` + `94112_home_zip` + `94114_home_zip` + `94115_home_zip` + `94116_home_zip` +
#               `94117_home_zip` + `94118_home_zip` + `94121_home_zip` + `94122_home_zip` + `94123_home_zip` +
#               `94124_home_zip` + `94127_home_zip` + `94131_home_zip` + `94132_home_zip` + `94133_home_zip` +
#               `94134_home_zip` + `94608_home_zip` + `94610_home_zip` + `94611_home_zip` + `94941_home_zip` +
#               `NULL_home_zip` + `other_home_zip`  + Cantonese + Mandarin + Spanish +
#               bachelor*one_jobdesc + bachelor*one_district_year + bachelor*two_district_year + bachelor*three_district_year +
#               bachelor*four_district_year + master*one_jobdesc + master*one_district_year + master*two_district_year + 
#               master*three_district_year + master*four_district_year + 
#               one_jobdesc*one_district_year + one_jobdesc*two_district_year + one_jobdesc*three_district_year +
#               one_jobdesc*four_district_year + male*one_district_year + male*two_district_year + male*three_district_year +
#               male*four_district_year + FTE_dummy*one_jobdesc + FTE_dummy*one_district_year + FTE_dummy*two_district_year + FTE_dummy*three_district_year +
#               FTE_dummy*four_district_year + `White/Caucasian`*one_jobdesc + `White/Caucasian`*one_district_year + `White/Caucasian`*two_district_year +
#               `White/Caucasian`*three_district_year + `White/Caucasian`*four_district_year + `White/Caucasian`*FTE_dummy + FTE_dummy*male + 
#               `White/Caucasian`*male + age_under_30 + age_30_40 + age_over_40 + age_under_30*male + age_30_40*male + age_over_40*male, 
#             data=training_set, family = binomial)
# 
# mod_full_rmse <-  rmse(full, test_set)
# 
# predicted_scores_full <- predict(full, type = "response")
# 
# pred_full <- prediction(predicted_scores_full, training_set$resignation)
# perf_full <- performance(pred_full, measure = "tpr", x.measure = "fpr")
# auc_full <- performance(pred_full, measure = "auc")@y.values[[1]]
# 
# 
# # Fit first forward model 
# Output <- step(null, scope=list(lower=null, upper=full), direction="forward")
# summary(Output)
                 
# slightly adjusted model with year variables removed to allow testing against test set
model1 <- glm(formula = resignation ~ one_jobdesc + 
    teacher + `White/Caucasian` + above_3_district_year +
    age_30_40 + master + one_n_licenses + bachelor + FTE_dummy + 
    age_under_30 + `94114_home_zip` + `94132_home_zip` + 
    `94123_home_zip` + `94117_home_zip` + `94118_home_zip` + 
    `94112_home_zip` + 
    `White/Caucasian` + FTE_dummy*`White/Caucasian` +
      teacher*master + teacher*bachelor + teacher*`White/Caucasian` + teacher*above_3_district_year, 
    family = binomial, data = training_set)


## assess model1


mod1_rmse <-  rmse(model1, test_set)

mod1_rmse

# predicted_scores <- predict(model1, type = "response")
# 
# 
# pred <- prediction(predicted_scores, training_set$resignation)
# perf <- performance(pred, measure = "tpr", x.measure = "fpr")
# plot(perf)
# abline(0,1,lty = 2)
# 
# 
# auc <- performance(pred, measure = "auc")@y.values[[1]]


# simplified model (model2)
model2 <- glm(formula = resignation ~ one_jobdesc + 
    teacher + `White/Caucasian` + above_3_district_year +
    age_30_40 + master + one_n_licenses + bachelor + FTE_dummy + 
    age_under_30,
    family = binomial, data = training_set)

## assess model2


mod2_rmse <-  rmse(model2, test_set)

mod2_rmse

mod2_rmse_holdout <- rmse(model2, holdout_set)

predicted_scores2 <- predict(model2, type = "response")


pred2 <- prediction(predicted_scores2, training_set$resignation)
perf2 <- performance(pred2, measure = "tpr", x.measure = "fpr")
plot(perf2)
abline(0,1,lty = 2)


auc2 <- performance(pred2, measure = "auc")@y.values[[1]]
```

Regress and round
```{r}
## get re-scaled coefficients for model2 (regress and round)

rescale_coefs <- function(coefficient, M = 10, estimates){
  
  new_coef <- coefficient* ( M/ (abs(max(estimates)) ) )
  
  return(round(new_coef,0))
  
}


model2_results <- tidy(model2)
model2_estimates <- model2_results$estimate
model2_varnames <- tbl_df(model2_results$term)

mod2_rescaled_coefs <- tbl_df(sapply(model2_estimates, rescale_coefs, estimate = model2_estimates))

final_reg_round_table2 <- bind_cols(model2_varnames, mod2_rescaled_coefs)



# distribution of risk scores in training set

risk_score <- training_set %>%
  mutate(risk = (9*one_jobdesc) +
                        (8*teacher) +
                        (5*`White/Caucasian`) +
                        (9*above_3_district_year) +
                        (6*age_30_40) +
                        (5*master) +
                        (5*one_n_licenses) +
                        (5*bachelor) +
                        (10*FTE_dummy)) %>%
  select(c(one_jobdesc, teacher, `White/Caucasian`, above_3_district_year, age_30_40, master, one_n_licenses, bachelor, FTE_dummy, risk, resignation))

hist(risk_score$risk[risk_score$resignation ==1])
				
summary(risk_score$risk[risk_score$resignation ==0])

ggplot(risk_score, aes(risk, fill = as.factor(resignation), color = as.factor(resignation))) +
  geom_density(na.rm = TRUE, alpha = .5) +
  labs(x = "Risk Score", y = "Density", title = "Distribution of Risk Scores", fill = "Resignation") + 
  theme_minimal()

```

